Superphy: Real-time prediction of \textit{E. coli} phenotype using graph-based genome sequence storage and retrieval with the \textit{spfy} module

### 2017-11-06 paper-webserver.tex
EDIT: some typing + merged the comparisons section

% 2. Why solving it is important (for Public health / research)
- everything is being sequenced (surveillance / outbreak / research)
- previously mentioned common analyses / want to leverage pre-computed results
- WGS does not discard old methods, linkage to thousands of historical results by developing in-silico methods of traditional tests
- need fast / standardized outbreak response
- need fast / standardized in-silico reference lab
- need fast / standardized storage and retrieval of results based on ontology
- all known data can be leveraged, allows the most informed decisions possible
- etc.

% 3. how we solved it
In this study, we present an update to the SuperPhy platform, called Spfy. Spfy enables large-scale genome processing in real-time and uniquely decomposses results into graph datastructures for further comparative genomics. Unlike traditional approaches which store results in separate data files or in a relational database, our graph datastructures are then integrated into a knowledge graph identifying relationships between all genomes sequenced in the past, present, and future. We present the use of graph databases as a solution to storing genome analysis results in a meaningful manner, and to allow retrospective comparisons as more genomes are sequenced or populations change.

@@

All results are stored as a series of linked nodes which enables the platform to build associations between results as they are generated. \par

@@

The entire Spfy update is packaged as a series of Docker \url{https://www.docker.com/} containers and connected together using Docker Compose \url{https://docs.docker.com/compose/}.

@@

% namely galaxy
Other scientific workflow technologies such as Galaxy \cite{goecks2010galaxy}, Kepler \cite{ludascher2006scientific}, and Taverna \cite{oinn2004taverna} have been applied to bioinformatic tasks.
Galaxy aims to provide a reproducible compuation-based research environment which is accessible to individuals without programming knowledge.
TODO: BAP and IRIDA, since they were the examples from above
Taverna ...
Kepler ...
To approach the challenge of integrating different bioinformatics programs, Spfy instead uses technologies prevalent in common web services not necessarily related to scientific workflows.
While we prefer to package individual bioinformatics programs using language-agnostic package managers such as Conda \url{https://conda.io/}, we do agree that suites of bioinformatics software benefit from having all dependencies installed and deployable in Docker containers \cite{di2015impact}.
These programs are connected to form a workflow using task queues; a design methodology used to offload long-running or non-instantaneous tasks over available computing infrastructure.
Unlike platforms such as Galaxy which favor a drag-and-drop design for building workflows, Spfy favors a tightly coupled approach explicitedy linking together different bioinformatics programs within the task queue implemntation.
In particular, Spfy uses asynchronous tasks queues which are tripped in response to user requests, thereby allowing immediate responsivity to users on the website, and processes generating human-readable results can be separated from processes handling long-term result storage.

@@

\subsection{Comparison with similar bioinformatic pipelines}

% para: \cite{naccache2014cloud}
% must install a ton of deps and even create dbs manually https://github.com/chiulab/surpi
% im not actually sure where they're getting the 'cloud-compatible' aspect of their paper from - its cloud compatible in that you deploy in the same manner as you would to bare-metal, but that's just like saying it 'runs on a computer'

% para: \cite{aanensen2016whole}
% makes some pretty pictures, but not really a pipeline

% para: \cite{joensen2014real,thomsen2016bacterial}
% we have a lot of similar directions to this, but just try to make it prettier, more friendly to use, and faster

TODO: The needs to be linked to the previous discussion above, as its absence was conspicuous.
The Bacterial Analysis Platform (BAP), developed out of the Technical University of Denmark, is the closest analogue to Spfy, and provides an integrated analysis pipeline for bacterial WGS data as a web service.
BAP is novel in its combined approach to genome analysis as different programs, such as for VF and AMR determination, are included by default into the pipeline; Spfy provides similar functionallity with an expanded focus on integrated result storage and big-data analyses.
In place of a MySQL database for storing the location of result files, Spfy parses result data into graph objects and integrates results into a persistant graph database.
This enables Spfy to create large datasets where analysis results are linked to genomes, and to perform population-wide analyses.
On a per file basis, Spfy performs at a similar speed to BAP on predictive genomics tasks, though Spfy does not provide genome assembly services.
Spfy also processes XXXX files over XX tasks in XXXX time by using a novel approach of distributing computations over a task queue and multiple Docker compartmentalized containers.

>>>>

% 2. Why solving it is important (for Public health / research)
WGS is becoming the norm for bacterial surveillance, outbreak tracing, and research \cite{ronholm2016navigating}. While pipelines exist \cite{joensen2014real} for running a variety of analyses on WGS data, 
In order to  Existing annotation and analysis software for WGS data represent units of compuatation; we can optimize WGS pipelines if these results are stored to avoid recomputation, and extend results by viewing them in the context of other analysis modules and genomes.
 these computations can be 
- everything is being sequenced (surveillance / outbreak / research)
- previously mentioned common analyses / want to leverage pre-computed results
- WGS does not discard old methods, linkage to thousands of historical results by developing in-silico methods of traditional tests
- need fast / standardized outbreak response
- need fast / standardized in-silico reference lab
- need fast / standardized storage and retrieval of results based on ontology
- all known data can be leveraged, allows the most informed decisions possible
- etc.

% 3. how we solved it
In this study, we present an update to the SuperPhy platform, called Spfy. Spfy enables large-scale genome processing in real-time and uniquely decomposses results into graph datastructures for further comparative genomics. Unlike traditional approaches which store results in separate data files or in a relational database, the graph datastructures are then integrated into a knowledge graph identifying relationships between all genomes sequenced in the past, present, and future. We present the use of graph databases as a solution to storing genome analysis results in a meaningful manner, and to allow retrospective comparisons as more genomes are sequenced or populations change.

@@

\par

@@

The entire Spfy update is packaged in Docker \url{https://www.docker.com/} containers and connected together using Docker Compose \url{https://docs.docker.com/compose/}.

@@

% namely galaxy
Other scientific workflow technologies such as Galaxy \cite{goecks2010galaxy}, and bioinformatics pipelines IRIDA and BAP are used to run analysis modules on WGS data.
Galaxy aims to provide a reproducible compuation-based research environment which is accessible to individuals without programming knowledge.
TODO: BAP and IRIDA, since they were the examples from above
IRIDA...
BAP...

The Bacterial Analysis Platform (BAP), developed out of the Technical University of Denmark, is the closest analogue to Spfy, and provides an integrated analysis pipeline for bacterial WGS data as a web service.
BAP is novel in its combined approach to genome analysis as different programs, such as for VF and AMR determination, are included by default into the pipeline; Spfy provides similar functionallity with an expanded focus on integrated result storage and big-data analyses.
In place of a MySQL database for storing the location of result files, Spfy parses result data into graph objects and integrates results into a persistant graph database.
This enables Spfy to create large datasets where analysis results are linked to genomes, and to perform population-wide analyses.
On a per file basis, Spfy performs at a similar speed to BAP on predictive genomics tasks, though Spfy does not provide genome assembly services.
Spfy also processes XXXX files over XX tasks in XXXX time by using a novel approach of distributing computations over a task queue and multiple Docker compartmentalized containers.

To approach the challenge of integrating different bioinformatics programs, Spfy instead uses technologies prevalent in common web services not necessarily related to scientific workflows.
While we prefer to package individual bioinformatics programs using language-agnostic package managers such as Conda \url{https://conda.io/}, we do agree that suites of bioinformatics software benefit from having all dependencies installed and deployable in Docker containers \cite{di2015impact}.
These programs are connected to form a workflow using task queues; a design methodology used to offload long-running or non-instantaneous tasks over available computing infrastructure.
Unlike platforms such as Galaxy which favor a drag-and-drop design for building workflows, Spfy favors a tightly coupled approach explicitedy linking together different bioinformatics programs within the task queue implemntation.
In particular, Spfy uses asynchronous tasks queues which are tripped in response to user requests, thereby allowing immediate responsivity to users on the website, and processes generating human-readable results can be separated from processes handling long-term result storage.

@@
